Nested sampling with two objective functions

### 2015-11-23 Paper/ms.tex
algorithm

We seek an algorithm for computing the partition function
$Z(\beta_1, \beta_2)$ for a range of values of the $\beta$s, from a single
run. To be considered a variant of Nested Sampling, the method should satisfy
the following properties:
\begin{enumerate}
\item The algorithm should begin with $N$ points drawn from the prior $\pi(\xx)$.
\item The algorithm should seek to explore regions where the values of
$L_1(\xx)$ and $L_2(\xx)$ are higher than the prior would typically predict.
\item The algorithm should consider a sequence of probability
distributions proportional to $\pi$, but restricted to smaller and smaller
domains for which we can approximately measure the enclosed prior mass.
\item The algorithm should be invariant to monotonic transformations of
$L_1$ and $L_2$, i.e. it should only depend on rankings of $L_1$ and $L_2$
values and not the values themselves.
\end{enumerate}
The standard Nested Sampling algorithm, (as well as alternatives such as
Diffusive Nested Sampling \citep{dnest}) has these properties.

@@

\begin{algorithm}
\begin{algorithmic}
\State Generate $N$ particles $\{\xx_1, ..., \xx_N\}$ from $\pi(\xx)$, and
initialise the {\em remaining prior mass} at $M=1$.
\While{more iterations desired}
	\State Calculate the {\it corner count} (Equation~\ref{eqn:corner_count})
of each particle.
	\State Choose one of the particles which has corner count equal to the
minimum nonzero value. Label this particle number $K$.
	\State Choose one of the objective functions at random. Label this
objective function $J$. Let $L_{J}(\xx_K)$ be the value of this objective
function for particle $K$.
\EndWhile
\end{algorithmic}
\end{algorithm}

Step 1 of the algorithm ensures that NS property 1 is trivially satisfied.
In step 2, we try to find a particle with the lowest possible nonzero
corner count. Often this will be 1. In Figure~\ref{fig:joint2}, particle $K$
is located at $(2.65, 1.31)$ and has a corner count of 1.
If all particles have a corner count of 0, the algorithm terminates.

>>>>

We seek an algorithm for computing the partition function
$Z(\beta_1, \beta_2)$ for a range of values of the $\beta$s, from a single
run. To be considered a variant of Nested Sampling, the method should satisfy
the following properties:
\begin{enumerate}
\item The algorithm should begin with $N$ points drawn from the prior $\pi(\xx)$.
\item The algorithm should seek to explore regions where the values of
$L_1(\xx)$ and $L_2(\xx)$ are higher than the prior would typically predict.
\item The algorithm should consider a sequence of probability
distributions proportional to $\pi$, but restricted to smaller and smaller
domains for which the enclosed prior mass shrinks by a (roughly) constant and
known factor each iteration.
\item The algorithm should be invariant to monotonic transformations of
$L_1$ and $L_2$, i.e. it should only depend on rankings of $L_1$ and $L_2$
values and not the values themselves.
\end{enumerate}
The standard Nested Sampling algorithm, (as well as alternatives such as
Diffusive Nested Sampling \citep{dnest}) has these properties.

@@

\begin{algorithm}
\begin{algorithmic}
\State Generate $N$ particles $\{\xx_1, ..., \xx_N\}$ from $\pi(\xx)$, and
initialise the {\em remaining prior mass} at $M=1$. The number of particles
$N$ should be odd and quite large (1001 or greater).
\While{more iterations desired}
	\State Calculate $\hat{X}(\ell_1, \ell_2)$ at the position of each particle
			and sort the $\hat{X}$ values observed.
	\State Find the median $\hat{X}$ value, $\hat{X}_{(k)}$, where $k = (N-1)/2$.
	\While{$k > 1$ and $\hat{X}_{(k-1)} = \hat{X}_{(k)}$}
		\State Set $k \leftarrow (k-1)$.
	\EndWhile
	\State Save all particles with $\hat{X} \geq \hat{X}_{(k)}$ to disk,
			assigning each a prior weight of $M/N$.
	\State Create a region in the $L_1$-$L_2$ plane defined by
				$\hat{X} \geq \hat{X}_{(k)}$, to be excluded at all subsequent
				steps.
	\State Update the remaining prior mass to $M \leftarrow (M - M/N)$.
	\State Replace all particles with $\hat{X} \geq \hat{X}_{(k)}$ by new
			particles generated from $\pi(\xx)$, subject to the constraint
			of not being in the excluded region.
\EndWhile
\end{algorithmic}
\end{algorithm}

Step 1 of the algorithm ensures that NS property 1 is trivially satisfied.
